Water Desalination Using Polyelectrolyte Hydrogel: Gibbs~Ensemble Modeling

### 2024-07-15 MDPI/proof/manuscript.v5/edit.tex
Cleaned up files. Article moved to MDPI directory

%=================================================================
% MDPI internal commands
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2022}
\copyrightyear{2022}
\externaleditor{Academic Editor: \hl{Firstname Lastname}} %MDPI: please add the Academic Editor
\datereceived{9 August 2022} 
%\daterevised{} % Only for the journal Acoustics
\dateaccepted{10 October 2022} 
\datepublished{} 
%\datecorrected{} % Corrected papers include a "Corrected: XXX" date in the original paper.
%\dateretracted{} % Corrected papers include a "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/} % If needed use \linebreak
%\doinum{}
%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

@@

\section{\highlighting{Model and Methods} %MDPI: to AE: please note that the section title in Gels Template: 1. Introduction; 2. Results and Discussion; 3. Conclusions; 4. Materials and Methods
\label{sec: theory}}

% water desalination technologies and Gibbs ensemble (https://doi.org/10.3390/nano10020293 \cite{erdos2020}).

@@

	\caption{The hydrogel compressed in \emph{open} and in \emph{closed} system.
		Red and blue circles represent $\na$ and$\cl$ ions. $V_0$ is the volume which gel has in free swelling equilibrium state.\label{fig:open and closed}}
\end{figure}
%In the case of \emph{open system} the gel is the grand canonical ensemle, as $\mu, V, T = \mbox{Const}$, while the case of \emph{closed system} the gel and the explicit reservoir are a variation of Gibbs ensemble \cite{Panagiotopoulos1988b} performed with constant volume for the both boxes, equal chemical potentials of ion species at constant temperature.

%Unfortunately, there is no universal computationally effective technique to sample pressure difference, volume, and number of particles at the same. 
Mechanical movement and the exchange of ions occur simultaneously in reality; however, we simulate them as alternating in a stop--run mode. 
To sample mechanical properties of the gel and reservoir, we use MD simulation, whereas to sample the ion distribution between the gel and a reservoir, we use MC simulation.
The details of this hybrid MCMD computational technique can be found in our previous studies of polyelectrolytes in open systems \cite{Rud2020, Rud2022, Landsgesel2020a}.

\includegraphics[width=0.7\textwidth]{{figures/simgibbs}.pdf}\caption{Diamond-like network in the simulation box. Color code represents
the individual ion types (red: $\na$, blue: $\cl$)
and the hydrogel (gray: neutral segment ($\AH$), cyan: charged segment
($\A$)). \label{fig:diamond}}
\end{figure}
Each monomer unit of the network carries a negative elementary electric charge.
Except for the gel monomers, the monovalent co- and counter-ions, $\cl$ and $\na$, are present in the simulation box.
The total electric charge of all the particles in the box is zero; therefore the number of $\na$ ions exceeds the number of $\cl$ ions by $\Ngel$.

\subsection{Algorithm}
As mentioned above, the whole simulation run consists of MD and MC subsimulations of mechanical movement of the particles and ion exchange. The algorithm follows:
\begin{enumerate}
\item Initiate the systems to simulate: the gel of volume $\Vgel$ and the external solution of volume $\Vout$;
\item Equilibrate the system, interspersing the MD and MC stages;
\item Run the MD subsimulation and collect the observables: pressure in both volumes, $P_\mathrm{gel}$, $P_\mathrm{out}$, and distances between the nodes of the gel network. The latter is needed to estimate the autocorrelation of the MD simulation;
\item Run the MC procedure, simulating ion exchange, and collect the number of ions in both boxes, 
$\Ncl\gel$ and $\Ncl\out$; 
\item Repeat the MD and MC subsimulations until the desired length of sample arrays is reached.
\end{enumerate}

@@

\section{Results and Discussion}

@@

\section{Conclusions}
\textls[-5]{We have modeled compression of a polyelectrolyte gel in thermodynamic equilibrium with a supernatant aqueous solution of limited amount.
We have shown that compression of the gel decreases the supernatant salinity.
We employed this phenomenon to model water desalination.
The desalination was done as a sequential combination of two processes:
(1) swelling of the gel in an \emph{open system}, exchanging ions with a large reservoir at constant salinity;
(2) compression of the gel in a \emph{closed system}, during which the gel exchanges ions with a small reservoir, affecting its salinity.
We estimated the energy consumption needed for producing one liter of potable water from brine and have shown that the proposed gel compression method may compete with modern desalination~technologies. %MDPI: we noticed that the supplementary file has been uploaded in submitting system. please confirm if there is a supplementary, if so, please cite it in main text and add section of supplementary materils “\supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.} ”. Please note that the reference in supplementary should be also included in mian text.
% Oleg Rud: There is no need of supplementary materials anymore, since we added the method description into the main text.
}

>>>>

%=================================================================
% MDPI internal commands
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2022}
\copyrightyear{2022}
%\externaleditor{Academic Editor: \hl{Firstname Lastname}} %MDPI: please add the Academic Editor
\datereceived{9 August 2022} 
%\daterevised{} % Only for the journal Acoustics
\dateaccepted{10 October 2022} 
\datepublished{} 
%\datecorrected{} % Corrected papers include a "Corrected: XXX" date in the original paper.
%\dateretracted{} % Corrected papers include a "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/} % If needed use \linebreak
%\doinum{}
%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

@@

%\section{\highlighting{Model and Methods} %MDPI: to AE: please note that the section title in Gels Template: 1. Introduction; 2. Results and Discussion; 3. Conclusions; 4. Materials and Methods
%\label{sec: theory}}
\section{Results and Discussion\label{sec: RD}}
% water desalination technologies and Gibbs ensemble (https://doi.org/10.3390/nano10020293 \cite{erdos2020}).

@@

	\caption{The hydrogel compressed in \emph{open} and in \emph{closed} system.
		Red and blue circles represent $\na$ and $\cl$ ions. $V_0$ is the volume which gel has in free swelling equilibrium state.\label{fig:open and closed}}
\end{figure}
%In the case of \emph{open system} the gel is the grand canonical ensemle, as $\mu, V, T = \mbox{Const}$, while the case of \emph{closed system} the gel and the explicit reservoir are a variation of Gibbs ensemble \cite{Panagiotopoulos1988b} performed with constant volume for the both boxes, equal chemical potentials of ion species at constant temperature.

%Unfortunately, there is no universal computationally effective technique to sample pressure difference, volume, and number of particles at the same. 
Mechanical movement and the exchange of ions occur simultaneously in reality; however, we simulate them as alternating in a stop--run mode. 
To sample mechanical properties of the gel and reservoir, we use Molecular Dynamics (MD) simulation, whereas to sample the ion distribution between the gel and a reservoir, we use Monte Carlo (MC) simulation.
The details of this hybrid MCMD computational technique can be found in our previous studies of polyelectrolytes in open systems \cite{Rud2020, Rud2022, Landsgesel2020a} and in Section~\ref{sec: MM}.

@@

@@

\section{Conclusions}
\textls[-5]{We have modeled compression of a polyelectrolyte gel in thermodynamic equilibrium with a supernatant aqueous solution of limited amount.
We have shown that compression of the gel decreases the supernatant salinity.
We employed this phenomenon to model water desalination.
The desalination was done as a sequential combination of two processes:
(1) swelling of the gel in an \emph{open system}, exchanging ions with a large reservoir at constant salinity;
(2) compression of the gel in a \emph{closed system}, during which the gel exchanges ions with a small reservoir, affecting its salinity.
We estimated the energy consumption needed for producing one liter of potable water from brine and have shown that the proposed gel compression method may compete with modern desalination~technologies. %MDPI: we noticed that the supplementary file has been uploaded in submitting system. please confirm if there is a supplementary, if so, please cite it in main text and add section of supplementary materils “\supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.} ”. Please note that the reference in supplementary should be also included in mian text.
% Oleg Rud: There is no need of supplementary materials anymore, since we added the method description into the main text.
}
\section{Materials and Methods\label{sec: MM}}

\includegraphics[width=0.7\textwidth]{{figures/simgibbs}.pdf}
\caption{
Diamond-like network in the simulation box. 
Red and blue particles are the ions $\na$ and $\cl$. \label{fig:diamond}}
\end{figure}
Each monomer unit of the network carries a negative elementary electric charge.
Except for the gel monomers, the monovalent co- and counter-ions, $\cl$ and $\na$, are present in the simulation box.
The total electric charge of all the particles in the box is zero; therefore the number of $\na$ ions exceeds the number of $\cl$ ions by $\Ngel$.

\subsection{Algorithm}
As mentioned above, the whole simulation run consists of MD and MC sub-simulations of the mechanical movement of the particles and ion exchange. The algorithm is the following:
\begin{enumerate}
\item Initiate the systems to simulate: the gel of volume $\Vgel$ and the external solution of volume $\Vout$;
\item Equilibrate the system, interspersing the MD and MC stages;
\item Run the MD subsimulation and collect the observables: pressure in both volumes, $P_\mathrm{gel}$, $P_\mathrm{out}$, and distances between the nodes of the gel network. The latter is needed to estimate the autocorrelation of the MD simulation;
\item Run the MC procedure, simulating ion exchange, and collect the number of ions in both boxes, 
$\Ncl\gel$ and $\Ncl\out$; 
\item Repeat the MD and MC subsimulations until the desired length of sample arrays is reached.
\end{enumerate}

### 2024-07-15 MDPI/proof/manuscript.v5/orig.tex
Cleaned up files. Article moved to MDPI directory

>>>>

\begin{table}[H]
\caption{Estimates of  %MDPI: Please confirm if the format of the table should be retained, e.g., the color of the font, the text direction, vertical line
% Oleg Rud: Confirm the colors and the directions of the table must retain
desalination efficiency. All units are calculated per one mol of gel segments. Values in brackets correspond to crossection of `red' and `orange' lines on the plots in Figures \ref{fig: PV and CV} and \ref{fig: NV and CN}.
\label{tab: table}}

\begin{adjustwidth}{-\extralength}{0cm}

		\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{1\fulllength}{@{\extracolsep{\fill}}ll|lc|c|l|ll}
\noalign{\hrule height 1pt}
\multicolumn{1}{l|}{\textbf{\boldmath{$c_{s}^{0}$}, mM}} & \textbf{\boldmath{$c_{s}^{5}$}, mM} & \multicolumn{1}{l|}{\textbf{\boldmath{$n_{\cl}^{0}$}, mM}} & \textbf{\boldmath{$n_{\cl}^{5}$}, mM} & \textbf{\boldmath{$\Delta \Vgel$}, L} & \textbf{\boldmath{$|W$}\textbar , J/L} & \multicolumn{2}{c}{\textbf{\boldmath{$W^{id}$}, J/L}}\tabularnewline

\noalign{\hrule height 0.5pt} 
\multirow{2}{*}{{\small{}$91.47$}} & \multirow{2}{*}{} & \multicolumn{2}{l|}{{\small{}$57.08\pm0.122\enskip\longrightarrow$}} & \multirow{2}{*}{{\small{}2.74}} & \multirow{2}{*}{{\small{}$95.4\pm1.9$}} &  & \multirow{2}{*}{}\tabularnewline
 &  & \multicolumn{2}{r|}{{\small{}$71.75\pm0.024$}} &  &  & \multirow{8}{*}{{\small{}\hspace{-1em}}\textcolor{teal}{\small{}$\left.\begin{array}{l}
\\
\\
\\
\\
\\
\\
\\
\\
\end{array}\right\rbrace $$\rotatebox{90}{\hspace{-5em}{\color{teal}\ensuremath{\begin{array}{cc}
W^{id}= & 52.9\\
W^{\mathrm{sim}}= & 202.8\pm3.2\\
R_{w}= & 0.54
\end{array}}}}$}} & \tabularnewline
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multicolumn{2}{l|}{{\small{}$89.41\pm0.23\enskip\longrightarrow$}} & \multirow{2}{*}{{\small{}56.90}} & \multirow{2}{*}{} & \multirow{2}{*}{{\small{}2.72}} & \multirow{2}{*}{{\small{}$109.1\pm1.7$}} &  & \multirow{2}{*}{}\tabularnewline
\multicolumn{2}{r|}{{\small{}$73.63\pm0.03$}} &  &  &  &  &  & \tabularnewline
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multirow{2}{*}{\textcolor{teal}{\small{}63.93}} & \multirow{2}{*}{} & \multicolumn{2}{l|}{\textcolor{teal}{\small{}$37.28\pm0.08\enskip\longrightarrow$}} & \multirow{2}{*}{\textcolor{teal}{\small{}3.26}} & \multirow{2}{*}{\textcolor{teal}{\small{}$100.9\pm1.7$}} &  & \tabularnewline
 &  & \multicolumn{2}{r|}{\textcolor{teal}{\small{}$50.579\pm0.013$}} &  &  &  & \multirow{8}{*}{{\small{}\hspace{-1em}}\textcolor{olive}{\small{}$\left.\begin{array}{l}
\\
\\
\\
\\
\\
\\
\\
\\
\end{array}\right\rbrace $$\rotatebox{90}{\hspace{-5em}\ensuremath{\begin{array}{cc}
W^{id}= & 38.2\\
W^{\mathrm{sim}}= & 207.3\pm2.2\\
R_{w}= & 0.53
\end{array}}}$}}\tabularnewline
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multicolumn{2}{l|}{\textcolor{teal}{\small{}$62.05\pm0.15\enskip\longrightarrow$}} & \multirow{2}{*}{\textcolor{teal}{\small{}37.17}} & \multirow{2}{*}{} & \multirow{2}{*}{\textcolor{teal}{\small{}3.18}} & \multirow{2}{*}{\textcolor{teal}{\small{}$107.4\pm1.4$}} &  & \tabularnewline
\multicolumn{2}{r|}{\textcolor{teal}{\small{}$48.21\pm0.02$}} &  &  &  &  &  & \tabularnewline
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multirow{2}{*}{\textcolor{olive}{\small{}44.91}} & \multirow{2}{*}{} & \multicolumn{2}{l|}{\textcolor{olive}{\small{}$23.75\pm0.06\enskip\longrightarrow$}} & \multirow{2}{*}{\textcolor{olive}{\small{}3.82}} & \multirow{2}{*}{\textcolor{olive}{\small{}$106.7\pm1.5$}} &  & \tabularnewline
 &  & \multicolumn{2}{r|}{\textcolor{olive}{\small{}$35.911\pm0.010$}} &  &  & \multirow{8}{*}{{\small{}\hspace{-1em}}\textcolor{red}{\small{}$\left.\begin{array}{l}
\\
\\
\\
\\
\\
\\
\\
\\
\\
\end{array}\right\rbrace \rotatebox{90}{\hspace{-5em}\ensuremath{\begin{array}{cc}
W^{\mathrm{id}}= & 43.8\ {\color{red}{\scriptstyle (41.0)}}\\
W^{\mathrm{sim}}= & 222.3\pm2.5\\
R_{w}= & 0.52\ {\scriptstyle (0.46)}
\end{array}}}$}} & \tabularnewline
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multicolumn{2}{l|}{\textcolor{olive}{\small{}$43.46\pm0.12\enskip\longrightarrow$}} & \multirow{2}{*}{\textcolor{olive}{\small{}24.27}} & \multirow{2}{*}{} & \multirow{2}{*}{\textcolor{olive}{\small{}3.91}} & \multirow{2}{*}{\textcolor{olive}{\small{}$106.4\pm1.1$}} &  & \tabularnewline
\multicolumn{2}{r|}{\textcolor{olive}{\small{}$30.795\pm0.008$}} &  &  &  &  &  & \tabularnewline
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multirow{2}{*}{\textcolor{red}{\small{}31.93}} & \multirow{2}{*}{} & \multicolumn{2}{l|}{\textcolor{red}{\small{}$14.66\pm0.04\enskip\longrightarrow$}} & \multirow{2}{*}{\textcolor{red}{\small{}4.20}} & \multirow{2}{*}{\textcolor{red}{\small{}$107.9\pm1.4$}} &  & \tabularnewline
 &  & \multicolumn{2}{r|}{\textcolor{red}{\small{}$25.297\pm0.006$}} &  &  &  & \multirow{8}{*}{{\small{}\hspace{-1em}}\textcolor{orange}{\small{}$\left.\begin{array}{l}
\\
\\
\\
\\
\\
\\
\\
\\
\\
\end{array}\right\rbrace \rotatebox{90}{\hspace{-5em}\ensuremath{\begin{array}{cc}
W^{id}= & 66.6\ {\scriptstyle (22.0)}\\
W^{\mathrm{sim}}= & 218.7\pm2.2\\
R_{w}= & 0.53\ {\scriptstyle (0.49)}
\end{array}}}$}}\tabularnewline
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multicolumn{2}{l|}{\textcolor{red}{\small{}$30.03\pm0.08\enskip\longrightarrow$}} & \multirow{2}{*}{\textcolor{red}{\small{}14.55}} & \multirow{2}{*}{} & \multirow{2}{*}{\textcolor{red}{\small{}$\begin{array}{c}
4.17\\
{\scriptstyle (3.22)}
\end{array}$}} & \multirow{2}{*}{\textcolor{red}{\small{}$\begin{array}{c}
115.6\pm0.9\\
{\scriptstyle (57.3)}
\end{array}$}} &  & \tabularnewline
\multicolumn{2}{r|}{\textcolor{red}{\small{}$\begin{array}{c}
18.585\pm0.004\,\\
{\scriptstyle (22.32)}
\end{array}$}} &  &  &  &  &  & \tabularnewline
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multirow{2}{*}{\textcolor{orange}{\small{}22.30}} & \multirow{2}{*}{} & \multicolumn{2}{l|}{\textcolor{orange}{\small{}$8.84\pm0.03\enskip\longrightarrow$}} & \multirow{2}{*}{\textcolor{orange}{\small{}$\begin{array}{c}
4.75\\
{\scriptstyle (3.76)}
\end{array}$}} & \multirow{2}{*}{\textcolor{orange}{\small{}$\begin{array}{c}
108.1\pm1.3\\
{\scriptstyle (68.3)}
\end{array}$}} &  & \tabularnewline
 &  & \multicolumn{2}{r|}{\textcolor{orange}{\small{}$\begin{array}{c}
\text{17.945}\pm\text{0.004}\\
{\scriptstyle (14.56)}
\end{array}$}} &  &  & \multirow{7}{*}{{\small{}\hspace{-1em}}\textcolor{magenta}{\small{}$\left.\begin{array}{l}
\\
\\
\\
\\
\\
\\
\\
\end{array}\right\rbrace \rotatebox{90}{\hspace{-5em}\ensuremath{\begin{array}{cc}
W^{id}= & 69.7\\
W^{\mathrm{sim}}= & 227.5\pm2.3\\
R_{w}= & 0.55
\end{array}}}$}} & \\[1ex]
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multicolumn{2}{l|}{\textcolor{orange}{\small{}$20.74\pm0.06\enskip\longrightarrow$}} & \multirow{2}{*}{\textcolor{orange}{\small{}8.83}} & \multirow{2}{*}{} & \multirow{2}{*}{\textcolor{orange}{\small{}$4.71$}} & \multirow{2}{*}{\textcolor{orange}{\small{}$110.8\pm0.8$}} &  & \tabularnewline
\multicolumn{2}{r|}{\textcolor{orange}{\small{}$11.082\pm0.002$}} &  &  &  &  &  & \tabularnewline
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multirow{2}{*}{\textcolor{magenta}{\small{}10.90}} & \multirow{2}{*}{} & \multicolumn{2}{l|}{\textcolor{magenta}{\small{}$3.00\pm0.01\enskip\longrightarrow$}} & \multirow{2}{*}{\textcolor{magenta}{\small{}6.08}} & \multirow{2}{*}{\textcolor{magenta}{\small{}$106.9\pm1.1$}} &  & \tabularnewline
 &  & \multicolumn{2}{r|}{\textcolor{magenta}{\small{}9.011$\pm$0.002}} &  &  &  & \tabularnewline
\cline{1-6} \cline{2-6} \cline{3-6} \cline{4-6} \cline{5-6} \cline{6-6} 
\multicolumn{2}{l|}{\textcolor{magenta}{\small{}$9.83\pm0.05\enskip\longrightarrow$}} & \multirow{2}{*}{\textcolor{magenta}{\small{}3.12}} & \multirow{2}{*}{} & \multirow{2}{*}{\textcolor{magenta}{\small{}5.78}} & \multirow{2}{*}{\textcolor{magenta}{\small{}$119.4\pm1.0$}} &  & \tabularnewline
\multicolumn{2}{r|}{\textcolor{magenta}{\small{}$3.862\pm0.001$}} &  &  &  &  &  & \\[2ex]
\noalign{\hrule height 1pt}
\end{tabularx}
\end{adjustwidth}
\end{table}
